%!TEX root = ../thesis.tex
\renewcommand{\chapterpath}{chapter2}
\renewcommand{\imgpath}{\chapterpath/img}

\chapter{Related Work}
\label{chapter:Introduction}
\minitoc

Why now: More and more intelligent system are coming into our daily life, making decision that can affect our everyday life. 

Why the readers: Interacting with such system is therefore more and more difficult and often requires the user to adapt to the machine.

Why me: We need to develop system where interaction is made easy and adaptive wrt. each particular users. 

Why this chapter: present the further logical content of the document. Present the general problem of learning from unlabeled instruction in general terms, just the challenges that is interesting 

What: We first give the general scientific and societal context of our work from which arises some question, and follow the general ideas that has evolved in our respective. We are going down from the big picture to the tiny corner of interest for us. we then provide the logical content of the document t answer the so what and what now question.

So what: we need to study this corner into more details

What now: let look at what others have done to tune our understanding of the problem and broaden our view

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Intelligent systems around}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Challenges}

Autonomy, long term and open-ended

\subsection{Sensing}

\subsection{Acting and planning}

\subsection{Social interaction}

\subsection{Design, morphology and appearance}

\subsection{Adapt, adapt, adapt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Approaches}

\subsection{Learning from human defined objective function: extrinsic motivation}

\subsection{Self-directed learning: intrinsic motivation}

\subsection{Learning from other: social interaction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Robot Learning from Humans}

\subsection{Learning from demonstrations}

\subsection{Learning from advises}

\subsection{Learning from reinforcement signals}

\subsection{How people teach robots}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Interactive Learning}

\subsection{Ambiguous protocols}

\subsection{Ambiguous signals}

\subsection{Active learner}

\subsection{Active teacher}

\subsection{Unknown signals, unknown tasks}

