%!TEX root = ../thesis.tex

\begin{vcenterpage}
% \noindent\rule[2pt]{\textwidth}{0.5pt}
\begin{center}
{\large\textbf{Acknowledgments\\}}
\end{center}


% It was all about people providing me with the opportunity to learn.

\noindent First of all, I would like to thank Pierre-Yves Oudeyer and Manuel Lopes for supervising me during these three years. Apart from their excellent advice and their open minded and supportive supervision, I particularly enjoyed the opportunity I was given to learn and to simply belong to the lab, listening and taking part in various enlightening scientific and philosophical discussions.\\

\noindent I am grateful to Dr. Charles Capaday, Dr. Ram\'{o}n Huerta and Prof. Auke Jan Ijspeert for providing me with the early opportunity to discover research in their respective labs. Without them, I would never have pursued doctoral studies.\\

\noindent During these three years, I had the chance to engage in different collaborative works. I thank I{\~n}aki Iturrate and Luis Montesano for their numerous advice and their infinite patience. I thank Anna-Lisa Vollmer and Katharina J. Rohlfing for their invaluable insight on the implication of this work to human behavioral understanding. I thank Peter Stone and Samuel Barrett for their enthusiasm and the opportunity I had to visit the \textsc{Larg} lab.\\

\noindent Many thanks to the members of my jury, Mich\`ele Sebag, Mohamed Chetouani, Luis Montesano and Fabien Lotte, for accepting to review this work. \\

% for they rich and insightful comments.\\

\noindent Numerous thanks to all the members of the \textsc{Flowers} team for contributing in making these three years a memorable experience: Matthieu, Pierre, Fabien, Olivier, Paul, Cl\'{e}ment, Thomas, Thomas, J\'{e}r\^{o}me, Haylee, Didier, Steve,  Mai, Thibault, Yoan, Benjamin, and Adrien. I also thank our team assistants Nicolas, Nathalie, Catherine, and the numerous internship students, especially the one I co-supervised: Mathieu, Axel, Julie, Chlo\'{e}, Brice, and Fabien.\\

\noindent Unlimited thanks to my family for the good start in life they gave me, my friends for they unique ability to make me laugh, and Sara for her support all along these three years.\\

% Without I{\~n}aki, running experiments involving brain signals would not have existed. He spent an uncountable amount of time sitting with the cup on and gel in the hair. Without Luis, my understanding of the problem would not . Him and Manuel were exceptionally patient when listening to my explanation. we would not have reached the same level of understanding of the problem.

% \noindent\rule[2pt]{\textwidth}{0.5pt}
\end{vcenterpage}